---
title: "redWine"
author: "Fernando Motta"
date: "February 28, 2018"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exploratory Data Analysis on Red Wine Composition by Fernando Motta
=======================================================================

```{r echo = TRUE, message = FALSE, packages}
# Loading the utilized libraries

library("ggplot2")
library("gridExtra")
library("corrplot")
library("magrittr")
library("pander")
library("memisc")
library("dplyr")
```

```{r echo = TRUE, Load_the_Data}
# Loading the dataset
data <- read.csv('wineQualityReds.csv')

```

## Univariate Plot Section

First, let us check some basic information and statistics about the data.


```{r echo = FALSE}
str(data)

summary(data)
```

This dataset has 13 variables and 1599 observables. This suggests an ample 
survey within the realm of port red wine.

Also, to become familiar with the dataset, I will plot the occurrence of the 
variables before analysing any relationship. To begin with, I will plot the 
distribution of wine quality.

```{r echo = FALSE}
ggplot(data = data, aes(x = quality)) + 
  geom_bar(width = 0.8, color = 'black', fill = I('green'))

```

It stands out from this plot that the range of the ratings is limited between 
3 and 8, which leads us to question whether the dataset is complete. Since these
are experienced people trying only a specific type of wine ("Vinho Verde") it 
seems likely that they are evaluating them with respect to every wine in
general. There is also a strong concentration in the ratings 5 and 6. It will be
interesting to observe whether the other characteristics appear in similarly
tight and sharp ranges.


```{r echo = FALSE}

#This is a brief explanation of all plot creations within this work. 
#The ggplot package is used to create the plots with the aesthetical parameters
#needed and the grid.arrange is used to organize them within the file


#Here, a boxplot is created
boxplot1 <- ggplot(data, aes(x = 1, y = fixed.acidity)) +
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
#              geom_jitter(alpha = 0.1) +
              geom_boxplot(alpha = 0.2, color = 'red') +
              scale_y_continuous(lim = c(4,16))

#Here, a histogram is created
hist1 <- ggplot(data = data, aes(x = fixed.acidity)) +
           geom_histogram(binwidth = 1, color = 'black', 
                          fill = I('orange')) +
           scale_x_continuous(lim = c(4,16))
#Here they are arranged and shown.
grid.arrange(boxplot1, hist1, ncol = 2)
```

This distribution is also very concentrated around 8, which is beginning to 
raise the suspicion that most "Vinho Verde" wines are somewhat similar.


```{r echo = FALSE}

#Similar code to what is documented for boxplot1
boxplot2 <- ggplot(data, aes( x = 1, y = volatile.acidity ) ) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
#               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               scale_y_continuous(lim = c(0,1.6))
hist2 <- ggplot(data = data, aes(x = volatile.acidity)) +
  geom_histogram(binwidth = 0.05, color = 'black',fill = I('orange')) +
  scale_x_continuous(lim = c(0,1.6))

grid.arrange(boxplot2, hist2, ncol = 2)
```

The volatile acidity presents a different characteristic from the ones observed 
so far, showing a bimodal profile. This distribution presents a few outliers,
but they don't affect the visualization so much, so I'll keep them.


```{r echo = FALSE}
#Similar code to what is documented for boxplot1
boxplot3 <- ggplot(data, aes( x = 1, y = citric.acid )) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
#               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' )
hist3 <- ggplot(data = data, aes(x = citric.acid)) +
  geom_histogram(binwidth = 0.08, color = 'black',fill = I('orange')) +
  scale_x_continuous(breaks = seq(0,1,0.1), lim = c(0,1))

grid.arrange(boxplot3, hist3, ncol = 2)
```

Citric acid shows a seemingly near random distribution, which can be attributed 
to the fact that it adds a characteristic to the flavour which is neither good 
nor bad (freshness), but dependent on the winemaker's goals.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot4 <- ggplot(data, aes( x = 1, y = residual.sugar )) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
              # geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               scale_y_continuous(lim = c(1,8))
hist4 <- ggplot(data = data, aes(x = residual.sugar)) +
  geom_histogram(binwidth = 0.1, color = 'black',fill = I('orange')) +
  scale_x_continuous(lim = c(1,8))

grid.arrange(boxplot4, hist4, ncol = 2)
```

This is another very skewed distributions, with a few outliers.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot5 <- ggplot(data, aes( x = 1, y = chlorides )) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
               #geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               scale_y_continuous(lim = c(0,0.25))
hist5 <- ggplot(data = data, aes(x = chlorides)) +
  geom_histogram(binwidth = 0.01, color = 'black',fill = I('orange')) +
  scale_x_continuous(lim = c(0,0.25))

grid.arrange(boxplot5, hist5, ncol = 2)
```

This is a distribution very similar to Residual Sugar.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot6 <- ggplot(data, aes( x = 1, y = free.sulfur.dioxide )) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
               #geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               scale_y_continuous(lim = c(0,45))
hist6 <- ggplot(data = data, aes(x = free.sulfur.dioxide)) +
  geom_histogram(binwidth = 1, color = 'black',fill = I('orange')) +
  scale_x_continuous(breaks = seq(0,80,5), lim = c(0,45))

grid.arrange(boxplot6, hist6, ncol = 2)
```

Once again we have a skewed distribution with a long tail and a high peak.
We also perceive a few outliers.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot7 <- ggplot(data, aes( x = 1, y = total.sulfur.dioxide )) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
               #geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               scale_y_continuous(lim = c(0,180))
hist7 <- ggplot(data = data, aes(x = total.sulfur.dioxide)) +
  geom_histogram(binwidth = 5, color = 'black',fill = I('orange')) +
  scale_x_continuous(lim = c(0,180))

grid.arrange(boxplot7, hist7, ncol = 2)
```

It was to be expected that this distribution would be very similar to the one 
found for free sulphur dioxide, since both quantities are related.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot8 <- ggplot(data, aes( x = 1, y = density)) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
               #geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' )
hist8 <- ggplot(data = data, aes(x = density)) +
  geom_histogram(binwidth = 0.001, color = 'black',fill = I('orange'))

grid.arrange(boxplot8, hist8, ncol = 2)
```

 
It is noteworthy that this is the first variable to exhibit a distribution that 
is very similar to the Normal Distribution. The same behaviour is observed for 
the pH distribution below.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot9 <- ggplot(data, aes( x = 1, y = pH)) + 
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +  
               #geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' )
hist9 <-ggplot(data = data, aes(x = pH)) +
              geom_histogram(binwidth = 0.1, color = 'black',fill = I('orange'))

grid.arrange(boxplot9, hist9, ncol = 2)

```


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot10 <- ggplot(data, aes( x = 1, y = sulphates)) + 
               #geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +
               scale_y_continuous(lim = c(0.3,1.6))
hist10 <- ggplot(data = data, aes(x = sulphates)) +
  geom_histogram(binwidth = 0.1, color = 'black',fill = I('orange')) +
  scale_x_continuous(lim = c(0.3,1.6))

grid.arrange(boxplot10, hist10, ncol = 2)

```

Sulphates present a behaviour we have come to expect, with a long tailed 
distribution, but it behaves better than the aforementioned distributions in the
sense that it has less outliers.


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Similar code to what is documented for boxplot1
boxplot11 <- ggplot(data, aes( x = 1, y = alcohol)) + 
               #geom_jitter(alpha = 0.1 ) +
               theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank()) +    
               geom_boxplot(alpha = 0.2, color = 'red' ) +
               scale_y_continuous(lim = c(8,14))
hist11 <- ggplot(data = data, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.1, color = 'black',fill = I('orange')) +
  scale_x_continuous(lim = c(8,14))

grid.arrange(boxplot11, hist11, ncol = 2)
```

Alcohol is once again similar, but it is not as skewed as the other
distributions.


## Analysis of the Univariate Plots

There seems to be a large concentration of wines along the central values of the
quality rating, which begs the question as to whether this dataset is complete.

It seems natural that pH might have a large influence on the taste, as well as
all acidity measures, since acid and basic substances have a characteristic 
taste. However, some of the present acids might have another flavouring 
characteristic which is more active than the drop in pH it causes. So it's 
important to observe the relationship between each acid and the pH and also
between them and the wine rating. 

The exception to this might be the citric acid, since its distribution has very 
particular characteristics.

Apart from citric acid, density and pH, the trend among the other variables 
seems to be a long tailed skewed distribution with some outliers.

## Bivariate Plots

To find out which relations are interesting, the first step would be to create a
correlation table which would ease visualization.

```{r}

#Creation of a correlation table from the data using pander
c <- cor(
  data %>%
    dplyr::select(-X))
#emphasize.strong.cells(which(abs(c) > .3 & c != 1, arr.ind = TRUE))
corrplot(c, method="color")
```


As water has a greater density than alcohol, it is in accordance to my 
expectations that Alcohol has a negative correlation with density. The variables
most strongly correlated with quality are Volatile acidity and alcohol. The 
relationship between Fixed acidity and density is also very strong.

To enhance this analysis, box plots are necessary, assessing each 
characteristic's association with quality, which is what we want to evaluate.


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = fixed.acidity)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

Fixed acidity does not seem to influence quality.


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = volatile.acidity)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

Volatile acidity seems to be inversely related to the quality of the wine, which
one might find surprising due to the correlation between quality and pH. On
further analysis, however, one will notice that Volatile acidity is positively
correlated with pH (therefore inversely correlated to acidity). 

This might mean that volatile acids are actually not kept in the wine, being 
lost to the air through their volatility and any are not in solution, but rather
dispersed.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = citric.acid)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```


The correlation between citric acid and quality is positive, which is unexpected
due to the citric acid distribution being near square and random. 

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = residual.sugar)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,5)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)


```


Residual sugar should not have a very large effect due to the nature of the 
winemaking process, where it exists as a byproduct of the fermentation. 
Therefore, the mean should be about the same along the range of qualities, which
is what we observe indeed.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = chlorides)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,0.2)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```

There seems to be a small correlation between the reduction of chloride and 
quality. However, I think this correlation is too small to be representative.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = free.sulfur.dioxide)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,40)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```

This seems to have an influence. Too low Free sulphur dioxide produces poor 
wine.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = total.sulfur.dioxide)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0,150)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```


This is expectedly similar to free sulphur, as it is the former's superset.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = density)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```


Lower density seems to result in better wine. This, however, might be related to
the lower density of alcohol.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = pH)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```


The pH seems to correlate inversely with the quality, which implies more acidic
wines are better.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = sulphates)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  scale_y_continuous(lim = c(0.25,1)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)


```

The correlation seems to be positive, which is normal because the ions are 
strong, therefore it corresponds with the pH observation.

```{r echo = FALSE, message = FALSE, warning = FALSE}

#Simply showing boxplots to visualize correlation with quality
ggplot(data = data, aes(x = factor(quality), y = alcohol)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)

```

Alcohol seems to be the main driver to wine quality. However, the amount of 
outliers suggests there may be another associated variable.

To evaluate this hypothesis, it's adequate to try to fit a linearisation between
alcohol and quality.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#This section creates a linear regression between alcohol and quality

alcoholQualityLinearisation <- lm(as.numeric(quality) ~ alcohol,
                       data = data)
summary(alcoholQualityLinearisation)
```

The low value observed for R squared implies that other variables also play a 
large role in the wine quality. To test this, I will make a correlation test 
between the variables and the quality individually.


```{r echo = FALSE, message = FALSE, warning = FALSE}

#This section calculates and shows the correlations between the variables.

simple_cor_test <- function(x, y) {
  return(cor.test(x, as.numeric(y))$estimate)
}

correlations <- c(
  simple_cor_test(data$fixed.acidity, data$quality),
  simple_cor_test(data$volatile.acidity, data$quality),
  simple_cor_test(data$citric.acid, data$quality),
  simple_cor_test(log10(data$residual.sugar), data$quality),
  simple_cor_test(log10(data$chlorides), data$quality),
  simple_cor_test(data$free.sulfur.dioxide, data$quality),
  simple_cor_test(data$total.sulfur.dioxide, data$quality),
  simple_cor_test(data$density, data$quality),
  simple_cor_test(data$pH, data$quality),
  simple_cor_test(log10(data$sulphates), data$quality),
  simple_cor_test(data$alcohol, data$quality))
names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'log10.residual.sugar',
                         'log10.chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'log10.sulphates', 'alcohol')

correlations
```

Some variables were analyzed through their logarithms due to their magnitude 
range. This test suggests that Alcohol, sulphates, volatile acidity and citric 
acid have a higher correlation with quality.

## Analysis of Bivariate Plots 

In addition to the effects with a positive correlation with quality noted above,
it is noteworthy that volatile acidity, chloride presence and density seem to 
have an inverse correlation, while fixed acidity and residual sugar have little
effect.

## Multivariate Plots

While alcohol is the main driver for quality, it still has a relatively small 
influence in the wine quality. This can be seen in the R squared factor. 
Therefore it is necessary to evaluate other factors. To do this we will keep 
alcohol constant and insert other variables. The variables chosen were the ones
shown in the former sections to have the most apparent impact.


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Creating plots to show combined influence from variables
#with different levels of alcohol. 
ggplot(data = data,
       aes(y = density, x = alcohol,
           color = as.factor(quality))) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size = 1)  +
  scale_color_brewer(type = 'seq',
                   guide = guide_legend(title = 'Quality'))
```

Density seems to be irrelevant given constant alcohol, confirming the causal 
relation between these factors.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Creating plots to show combined influence from variables
#with different levels of alcohol. 
ggplot(data = data,
       aes(y = sulphates, x = alcohol,
           color = as.factor(quality))) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size = 1)  +
  scale_y_continuous(limits = c(0.3,1.5)) +
 # facet_wrap(~rating) +
  scale_color_brewer(type = 'seq',
                   guide = guide_legend(title = 'Quality'))
```

It seems that, given constant alcohol, higher levels of sulphates produce better 
wine.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Creating plots to show combined influence from variables
#with different levels of alcohol. 
ggplot(data = data,
       aes(y = volatile.acidity, x = alcohol,
           color = as.factor(quality))) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
#  facet_wrap(~rating) +
  scale_color_brewer(type = 'seq',
                   guide = guide_legend(title = 'Quality'))
```

On the other hand, given constant alcohol, higher volatile acids produce poorer
wines.



```{r echo = FALSE, message = FALSE, warning = FALSE}
#Creating plots to show combined influence from variables
#with different levels of alcohol. 
ggplot(data = data,
       aes(y = pH, x = alcohol,
           color = as.factor(quality))) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size = 1)  +
 # facet_wrap(~rating) +
  scale_color_brewer(type = 'seq',
                   guide = guide_legend(title = 'Quality'))
```

Expectedly, lower pH with constant alcohol produces better wine. We had 
previously seen that volatile acidity and pH influence the wine in opposite
ways.

Given this data, it's interesting to generate a model considering the variables
which correlate more strongly with quality.


```{r}

#This block of code generates the data that will allow us to construct the
#table, which will then allow us to visualize which variables correlate more
#strongly with wine quality

set.seed(1056)
training_data <- sample_frac(data, .6)
test_data <- data[ !data$X %in% training_data$X, ]
m1 <- lm(as.numeric(quality) ~ alcohol, data = training_data)
m2 <- update(m1, ~ . + sulphates)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + fixed.acidity)
m6 <- update(m2, ~ . + pH)
mtable(m1,m2,m3,m4,m5,m6)
```


```{r echo = FALSE, message = FALSE, warning = FALSE}
wine_predict <- data.frame(
  test_data$quality,
  predict(m5, test_data) - as.numeric(test_data$quality)
)
names(wine_predict) <- c("quality", "error")
ggplot(data = wine_predict, aes(x = quality,y = error)) +
  geom_jitter(alpha = 0.3)
```

## Analysis of Multivariate Plots

The variables more strongly correlated with the quality of wine in combination 
seem to be alcohol. Sulphate and Citric acid. Also, the linear models presented 
low values for R squared, which suggests that the data might not be significant 
enough to draw definite conclusions and this study would profit from more 
extensive datasets.

## Final Plots

Given the conclusion that Alcohol and Sulphates affect quality most 
proeminently, I think that the most important graphics display these 
characteristics.


```{r echo = FALSE, message = FALSE, warning = FALSE}
#boxplot showing other statistical parameters within it
ggplot(data = data, aes(y = alcohol, x = quality)) + 
  #geom_jitter(alpha = .3)  +
  geom_boxplot(alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  xlab("Quality") +
  ggtitle("Influence of alcohol on wine quality")
```

This plot shows that higher alcohol percentages generate better wines. Since for
each box plot the mean and median concide, we infer that for each particular 
quality, the alcohol distribution is almost normal, which means that the high 
median in the top range implies that nearly every high quality wine has high 
alcohol rating. But, as we mentioned before, the low R squared means that other 
variables are also relevant.

```{r echo = FALSE, message = FALSE, warning = FALSE}
#Combined correlation of alcogol and sulphate
ggplot(data = data,
       aes(y = sulphates, x = alcohol,
           color = as.factor(quality))) +
  geom_point(alpha = 0.8, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size = 1)  +
  scale_y_continuous(limits = c(0.3,1.5)) +
  ylab("potassium sulphate (g/dm3)") +
  xlab("Alcohol Percentage") +
  scale_color_brewer(type = 'seq',
                   guide = guide_legend(title = 'Quality')) +
  ggtitle("Alcohol and sulphates over wine quality")
```

This plot indicates that high alcohol contents and high sulphate contents 
produce better wines in combination, which gives a more consistent
characteristic to search for in wines.


```{r}
#Analysis of the predictive model accuracy
df <- data.frame(
  test_data$quality,
  predict(m5, test_data) - as.numeric(test_data$quality)
)
names(df) <- c("quality", "error")
ggplot(data = df, aes(x = quality, y = error)) +
  #geom_jitter(alpha = 0.3) +
  ggtitle("Linear model errors vs expected quality")
```

This plot serves the purpose of illustrating the errors in the linearisation.
One can easily notice that the results are worse for the extreme ranges 
(poor wine and good wine), which can be attributed to the much higher amount of
average samples.

## Reflections

This dataset seems to have been treated beforehand, because it is extremely
well behaved. Having said that, my biggest difficulty in this analysis was
related to the fact that the distribution of wine quality was too centered 
around the average values, lacking data in the extremes.

This particularity in the dataset hindered the creation of a more accurate
model, in my opinion. Another problem I saw was the fact that no single 
variable correlated strongly with quality, but then again, this is why 
sophisticated data analysis techniques are developed and employed.

As a future work this analysis might be extended to a larger dataset, 
possibly including other regions and grapesm favouring a more complete 
dataset.